Bare minimum: set theory

### 2023-10-23 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: minor correction
SetTheory: using "connected" instead of "total" for relationships

\begin{defn}[Homogeneous relation properties]
	Let $R \subseteq A \times A$ \marginleft{Reflexivity, symmetry, transitivity, ...} be a homogeneous binary relation between $A$ and itself. We define the following properties:
	\begin{description}
		\item[Reflexive.] Every element $a \in A$ is related to itself. That is, $aRa$ for all $a \in A$.
		\item[Irreflexive.] No element $a \in A$ is related to itself. That is, $aRa$ for no $a \in A$.
		\item[Symmetric.] If $a$ is related to $b$, then $b$ is related to $a$. That is, if $aRb$ then $bRa$.
		\item[Antisymmetric.] If $a$ is related to $b$ and $b$ is related to $a$, then $a$ and $b$ are the same element. That is, if $aRb$ and $bRa$ then $a=b$.
		\item[Transitive.] If $a$ is related to $b$ and $b$ is related to $c$, then $a$ is related to $c$. That is, if $aRb$ and $bRc$ then $aRc$.
		\item[Total.] Given two distinct elements, one is related to the other.That is, given $a \neq b$, either $aRb$ or $bRa$. (This has no relation to left/right-total)
	\end{description}

@@

\begin{defn}[Orders and equivalence]
	A \textbf{partial order} \marginleft{Order, equivalence: $\leq$, $<$, $\equiv$} (noted for example $\leq$, $\lesssim$, $\preceq$)  is a homogeneous binary relationship that is reflexive, antisymmetric and transitive. A \textbf{linear order} or \textbf{total order} is an order that is total. An order is \textbf{strict} (noted for example $<$, $\prec$) if it is irreflexive instead of reflexive.

@@

\begin{prop}
	The following statements are equivalent:
	\begin{itemize}
		\item Axiom of choice: every set has a choice function
		\item Well-ordering theorem: every set can be well ordered (defined below)
		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other (i.e. $|A| < |B|$, $|A| = |B|$ or $|A| > |B|$)
		\item Zorn's lemma: every non-empty poset (i.e. partially ordered set) for which every chain (i.e. totally ordered subset) has an upper bound contains at least one maximal element
		\item Every vector space has a basis
	\end{itemize}
\end{prop}

>>>>

\begin{defn}[Homogeneous relation properties]
	Let $R \subseteq A \times A$ \marginleft{Reflexivity, symmetry, transitivity, ...} be a homogeneous binary relation between $A$ and itself. We define the following properties:
	\begin{description}
		\item[Reflexive.] Every element $a \in A$ is related to itself. That is, $aRa$ for all $a \in A$.
		\item[Irreflexive.] No element $a \in A$ is related to itself. That is, $aRa$ for no $a \in A$.
		\item[Symmetric.] If $a$ is related to $b$, then $b$ is related to $a$. That is, if $aRb$ then $bRa$.
		\item[Antisymmetric.] If $a$ is related to $b$ and $b$ is related to $a$, then $a$ and $b$ are the same element. That is, if $aRb$ and $bRa$ then $a=b$.
		\item[Transitive.] If $a$ is related to $b$ and $b$ is related to $c$, then $a$ is related to $c$. That is, if $aRb$ and $bRc$ then $aRc$.
		\item[Connected.] Given two distinct elements, one is related to the other. That is, given $a \neq b$, either $aRb$ or $bRa$.
	\end{description}

@@

\begin{defn}[Orders and equivalence]
	A \textbf{partial order} \marginleft{Order, equivalence: $\leq$, $<$, $\equiv$} (noted for example $\leq$, $\lesssim$, $\preceq$)  is a homogeneous binary relationship that is reflexive, antisymmetric and transitive. A \textbf{linear order} or \textbf{total order} is an order that is connected. An order is \textbf{strict} (noted for example $<$, $\prec$) if it is irreflexive instead of reflexive. A \textbf{strict linear order} or \textbf{strict total order} is an order that is strict and connected.

@@

\begin{prop}
	The following statements are equivalent:
	\begin{itemize}
		\item Axiom of choice: every set has a choice function
		\item Well-ordering theorem: every set can be well ordered (defined below)
		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other (i.e. $|A| < |B|$, $|A| = |B|$ or $|A| > |B|$)
		\item Zorn's lemma: every non-empty poset (i.e. partially ordered set) for which every chain (i.e. linearly ordered subset) has an upper bound contains at least one maximal element
		\item Every vector space has a basis
	\end{itemize}
\end{prop}
